---
title: "Making growth curve with data from eVOLVER"
author: "cmkim"
date: "2023-11-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

##### Loading library

```{r library}
library(readxl)
library(dplyr)
library(ggplot2)
```

##### Loading data
Data from eVOLVER is saved in .csv format.
Each value means optical value sensed by each sleeve, every 1 second.
"Seonsor.value" means raw value from sensor, "OD.value" means calibrated value of Sensor.value.
```{r loading_data}
data <- read.csv("/Users/chungminkim/5projects/08MB_interaction/MHB_ZGAL/od_save_9445.csv")
data <- as.data.frame(data)
head(data)
```

##### Input OD value measured at start point & end point of growth
Raw value is converted into OD value.
You can use calibrated data by changing column.
```{r growth data}
ODstart1 <- 0.042
ODfinish1 <- 0.705

ODstart2 <- 0.04
ODfinish2 <- 0.782

ODstart4 <- 0.044
ODfinish4 <- 0.678

n <- nrow(data)

growthS1 <- colMeans(data[(n-1200):n, 1, drop=F]) - colMeans(data[1:1200, 1, drop=F])
growthS2 <- colMeans(data[(n-1200):n, 2, drop=F]) - colMeans(data[1:1200, 2, drop=F])
growthS4 <- colMeans(data[(n-1200):n, 4, drop=F]) - colMeans(data[1:1200, 4, drop=F])

growth1 <- ODfinish1 - ODstart1
growth2 <- ODfinish2 - ODstart2
growth4 <- ODfinish4 - ODstart4

for(i in 1:n){
  data[i,5] <- (data[i,1] - data[1,1])*growth1/growthS1 + ODstart1
}
for(i in 1:n){
  data[i,6] <- (data[i,2] - data[1,2])*growth2/growthS2 + ODstart2
}
for(i in 1:n){
  data[i,8] <- (data[i,4] - data[1,4])*growth4/growthS4 + ODstart4
}
head(data[ ,5:8])
```

##### OD curve table
Making OD curve table which contains mean OD value sensed for 20 minutes (1,200 seconds).
```{r OD curve table}
ODcurve <- data.frame(OD=numeric(0), time=numeric(0), sleeve=character())
for(i in 1:(floor(n/1200)-1)){
  ODcurve[nrow(ODcurve)+1, ] <- c(colMeans(data[((i-1)*1200+1):(i*1200), 5, drop=F]),
                                  (1/3)*(i-1), "sleeve1")
}
for(i in 1:floor(n/1200)){
  ODcurve[nrow(ODcurve)+1, ] <- c(colMeans(data[((i-1)*1200+1):(i*1200), 6, drop=F]),
                                  (1/3)*(i-1), "sleeve2")
}
for(i in 1:floor(n/1200)){
  ODcurve[nrow(ODcurve)+1, ] <- c(colMeans(data[((i-1)*1200+1):(i*1200), 8, drop=F]),
                                  (1/3)*(i-1), "sleeve4")
}
ODcurve1 <- head(ODcurve, n=3)
ODcurve2 <- head(ODcurve[(floor(n/1200)):(2*floor(n/1200)),], n=3)
ODcurve3 <- head(ODcurve[(2*floor(n/1200)):(3*floor(n/1200)),], n=3)

rbind(ODcurve1, ODcurve2, ODcurve3)
```


##### OD plot table
To make Growth curve, calculate mean values and standard deviations of OD.
```{r Growth cruve table}
ODcurve$OD <- as.numeric(ODcurve$OD)
ODcurve$time <- as.numeric(ODcurve$time)
ODplot <- ODcurve %>% group_by(time) %>%
  summarize(OD_avg=mean(OD), OD_sd=sd(OD))
ODplot <-ODplot %>% arrange(time)

head(ODplot)
```

##### Making growth curve
```{r Growth cruve}
my_title <-  expression(paste(bold("Growth Curve of "), bolditalic("Z. galactanivoran"), bold( " Dsij")))
ODplot %>% 
#  filter(time<12.1) %>%
  ggplot(aes(x=time, y=OD_avg)) +
  geom_line() +
  geom_point(size=5) +
  #  geom_smooth() +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size=30, face="bold"),
        axis.title = element_text(hjust = 0.5, size=30, face="bold"),
        axis.text.x = element_text(size=30),
        axis.text.y = element_text(size=30),
        axis.ticks.length = unit(.3, "cm"),
        legend.text=element_text(size=30)) +
  labs(title = my_title, x = 'Time(hour)', y = 'OD600') +
  geom_errorbar(aes(x=time, ymin=OD_avg-OD_sd, ymax=OD_avg+OD_sd), width=0.3)
```